---
title: "The First Post is the Hardest"
author: "Renée Link"
date: 2020-08-02T00:09:12-08:00
categories: ["R"]
tags: ["critical pedagogy", "dpl", "labs", "specs grading", "remote teaching"]
draft: true
---

# Who am I, and why am I writing this?
> "You should write a blog."

> "Are you going to write this up somewhere?"

> "Do you have that written up somewhere that I can find it and share it?"

Over the span of about a week, I lost track of how many people asked me some variation of this. It came from Twitter, emails from acquaintances and strangers, and fellow educators in the [Digital Pedagogy Lab](https://www.digitalpedagogylab.com/dpl-2020-online/) (DPL) STEM-H track (more about that later) So I guess I'm writing a blog now!

Why would anyone be interested in what I have to say? Well, you see, I teach lab classes. Very, very large lab classes. Like, 1,000+ students and 30+ graduate teaching assistants large classes. While there are quite a few of us around the country (and probably the world), running the massive lab programs can be a very isolating job. There are usually only one or two of us at any given school, and it's often hard for us to find each other at different schools. Our positions can land anywhere on a spectrum from "regular" tenure-track faculty to contingent faculty. I've been trying to make connections with colleagues in similar positions for years with moderate success. We all face similar challenges. Perhaps putting myself out here more publicly will help facilitate more connections so we can help each other out.

# DPL and unleashing my superpower
Much of the prompting that led me to start writing here came from the DPL STEM-H track, so let me get back to that. My journey to DPL followed my usual MO. I'll describe it in those terms.

## Step 1: Frustration
I often see other educators talking about really cool things they can do with their classes and get really frustrated and a tad jealous. There are so many amazing things you can do with smaller class sizes that just become impossible when your enrollments hit quadruple digits. My original reaction to adcovates of critical pedagogy is a great example.

I came across [DPL on Twitter](https://twitter.com/DigPedLab). My initial response to many of the conversations around critical pedagogy and digital critical pedagogy was a combination of frustration, feeling left out, and feeling judged. All of these feelings stem from the fact that I rarely saw anyone talking about critical pedagogy in terms of STEM courses (except [Rissa Sorensen-Unruh](https://twitter.com/RissaChem) - go follow her now and thank me later). When I did see any STEM-related conversations, they always seemed to be centered around small-enrollment courses. This led to me feel left out of the conversation -- outside looking in. 

```{r echo=FALSE}
blogdown::shortcode("figure", src = "/post/first-post_files/outside_looking_in.jpg", caption = "[outside looking in](https://www.flickr.com/photos/bnilsen/8376023547) by bnilsen", attr = "outside looking in by bnilsen", attrlink = "https://www.flickr.com/photos/bnilsen/8376023547", alt = "image of cat looking in window")

```


And then there was the (perceived - probably reflecting my own insecurities) judgment. The message I was getting was that if you weren't co-creating your syllabus with your students, getting rid of assignment deadlines, or letting students decide what they wanted to learn, then you weren't doing enough to support your students. I can't count the number of times I yelled at my phone, "But HOW do I do that with a thousand students!?!" To their credit, anytime I decided to interact with a well-known person from this crowd and bring up a concern , the response I received was gracious and kind. Here's an example about trying to find ways to move organic chemistry labs online as we all went into panedemic lockdown.

```{r echo=FALSE}
blogdown::shortcode('tweet', '1239717886794457089')
blogdown::shortcode('tweet', '1239736309222199302')

```

So, yeah. The judgment was probably more in my head than in reality. I'll own that. When DPL2020 moved online, I decided to check it out. When I found the STEM-H track, I was all in. 

Our guide (Rissa) and my fellow DPL participants were lovely. We had thought-provoking conversation about how we could humanize teaching in STEM-H courses. Me being me, I repeatedly brought up the massive lab class perspective.

* How do I attempt to humanize online STEM teaching when much of the student contact is with graduate teaching assistants who may not buy into my pedagogical philosophy?
* If we adopt the stance that students should not have to pay for any resource, how do I use any digital tools in my classes? Free tools like Google Docs don't have the backend features I need when teaching at scale. How do I track over 1,000 different Google docs?
* How do I balance humanizing teaching for students without overburdening the teaching assistants, who are simultaneously staff *and* students?
* I want to trust students, but with this many, someone is inevitably copying someone else's post-lab assignment. Is using plagiarism detection software inevitable here? It's not realistic to change experiments from year to year in the way a humanities class can change an essay or project prompt.


## Step 2: My adaptation superpower
This what I'm good at. I see some amazing aspect of pedagogy that initially I think will be impossible to use at scale. I grumble to myself about it for a while. Then the creative, problem solving part of my brain takes over and It. Will. Not. Stop. Once I start to get an idea of how to adapt for scale, it's like my brain is on fire! After furious scribbling in one of my many notebooks, often over a span of hours, I come away with the outline of plan that adapts the spirit of the thing to the reality of my teaching situations. I've done this with so many things: large-scale flipped classes and specifications grading, to name a couple.

```{r echo=FALSE}
blogdown::shortcode("figure", src = "/post/first-post_files/IMG_20190206_221242.jpg", title = "scribbled notes on specifications grading", height = '50%', width = '50%')

```


Through DPL, I discovered that I had already been applying some aspects of critical pedagogy to my course design. A lot of this has come in lately since my adoption of specifications grading. I don't have the brilliant adaptations of additional aspects yet, but I'm working on it. That brain on fire moment is coming. The kindling has caught. The blaze won't be far behind.

## Step 3: Tell everyone I meet about it
Once I enact my plan, I can't stop talking about it. I tell anyone who will listen, and even those who don't want to. Just ask anyone who has interacted with me for more than five minutes. 

I also freely share every resource I have. This is where the blog comes in. I'm writing up much of what I adapt as academic publications, but that process takes time. Some people won't be able to access publications behind paywalls, so I do post all of mine as preprints on [ChemRxiv](https://chemrxiv.org/). But so much of what I've been adapting lately is urgently needed as we all grapple with remote teaching for the fall (and let's admit it - probably winter and spring, too). I have experience to share, and I feel that I have to make it available to everyone faster than the formal publishing process will allow.

# So what can you expect to see soon?
Here is what I already have available elsewhere.

* How we put labs online for spring in a hurry
  + [Preprint version](https://chemrxiv.org/articles/preprint/Online_in_No_Time_Design_and_Implementation_of_a_Remote_Learning_First_Quarter_General_Chemistry_Laboratory_and_Second_Quarter_Organic_Chemistry_Laboratory/12605207)
  + Published version - Accepted! Coming soon
  + [Public Google Drive](https://drive.google.com/drive/folders/1hpz1P3ejHDtjsTi-ZFSIvOR6yZ4GXxxr?usp=sharing) with all of our materials
* Specifications grading in lab small scale pilot
  + [Preprint version](https://chemrxiv.org/articles/preprint/Developing_and_Implementing_a_Specifications_Grading_System_in_an_Organic_Chemistry_Laboratory_Course/12276701)
* Some [specifications grading resources](https://faculty.sites.uci.edu/reneelink/specifications-grading-resources/) on my faculty website
* My [interview](https://www.alchem.ie/blog/specifications-grading-at-uc-irvine-with-renee-link) with Julia from Alchemie

What I'm planning to post here soon:

* How I converted my fully flipped, in-person, 5-week lecture course to remote delivery using specifications grading and no proctored exams
* How I used a mix of synchronous and asynchronous work in the lecture class, including managing group work in Zoom
* How I'm adapting lab TA training to a remote format and for remote lab classes
* How you can do specifications grading in massive classes
* How I'm planning for my winter quarter giant lab class in remote format
* My ideas for how to get students, especially chemistry majors, up to speed with hands-on lab techniques when we can eventually do things in person again
* Anything else people start asking me about on Twitter

I'll prioritize the first three items because I think they're of the most immediate use and interest to others.

Eventually I'll probably start writing about some other things like:

* Being first gen - it never ends, but you get to help others coming up after you
* Trying to transition from bench chemistry to CER-lite
* What we're doing next with labs (Hint: ADI!)
* Providing grad students opportunities to explore pedagogy, SoTL, and an intro to CER when we don't have an actual CER program
* Random dog pictures.

Let's go.

```{r echo=FALSE}
blogdown::shortcode("figure", src = "/post/first-post_files/dog-walking.jpg", title = "walking dogs")

```

